### 2013-07-29 in_progress/presentations/poster/reuMHC2013NoSIRIntro.tex

\documentclass[final]{beamer}
%Poster Packages: 
%dimensions are in cm, scale is for font size
\usepackage[size=custom,height=91.4,width=121.9,scale=1.8]{beamerposter}
%General Packages:
\usepackage{amsmath,amsthm, amssymb, latexsym}
%TikZ:
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\input{arrowsnew}

%Theme:
\mode<presentation>{\usetheme{I6pd2}}

\graphicspath{{figures/}}

%\newlength{\columnheight}
%\setlength{\columnheight}{88cm}

\title{\huge Modeling Intervention Strategies for United States TB Control}
\author{Jessica Ginepro, Emma Hartman, Ryo Kimura, Matthew McDermott, Colin
        Pawlowski, \& Dylan Shepardson}
\institute[MHC]{Mathematical Modeling Group, Mount Holyoke College, South
                Hadley, MA, USA}
\date[July 31, 2013]{July 31, 2013}
\begin{document}
%For the compartmental models:
\tikzstyle{compartment} = [rectangle, draw, fill=blue!20, text badly centered,
  node distance=3.8em, inner sep=0.4em, rounded corners]
\tikzstyle{compartment rate} = [right of=start, color=black, node distance=2.2em,
text width=3em]
%TODO: Increase arrow head size
\tikzstyle{line} = [draw, -latexnew, arrowhead = 7mm]

\begin{frame}
  \begin{columns}[T]
    \begin{column}{.3\textwidth}
      \begin{block}{Introduction}
        Epidemiological models offer insight into the structure of disease outbreaks and the merits of various interventions.  Compartmental differential equation models are a common model in which populations move between various health states, or compartments, according to predetermined rates.  This work is an extension of the Hill Model, a complex compartmental model of tuberculosis (TB) in the United States.
      \end{block}
      \vspace{1em}
      \begin{block}{The Basic Hill Model}
        \begin{columns}[T]
          \begin{column}{.5\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[width=\textwidth]{HillModelFlowChart}
              \end{center}
              \caption{The Hill Model schematic}
              \label{fig:hillFlow}
            \end{figure}
          \end{column}
          \begin{column}{.45\textwidth}
              \vspace{1.5em}
            Populations:
            \begin{itemize}
              \item US Born (USB) 
              \item Foreign Born (FB)
              %\item USB TB incidence declining
              %\item FB latent TB infection (LTBI) high
              %\item TB elimination in total population not projected by 2100
            \end{itemize}
            Individuals also leave the model due to natural death.
          \end{column}
        \end{columns}
        \vspace{0.6em}
        \begin{itemize}
          \item USB TB incidence rates are declining
          \item FB latent TB infection (LTBI) arrivals remain high
          \item TB elimination in total population not projected by 2100
          \item TB cases and costs dominated by FB LTBI individuals
        \end{itemize}
        \vspace{0,6em}
        \begin{columns}[T]
          \begin{column}{.47\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[height=0.9\textwidth,width=\textwidth] {incPlotSourced2}
              \end{center}
              \caption{Source population of US TB incidence}
              \label{fig:incPlotSourced}
            \end{figure}
          \end{column}
          \begin{column}{.47\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[height=0.9\textwidth,width=\textwidth]{costPlotSourced2}
              \end{center}
              \caption{Source population of US HCS TB cost}
              \label{fig:incPlotTotal}
             \end{figure}
           \end{column}
         \end{columns}
       \end{block}
    \end{column}

    \begin{column}{.3\textwidth}
      %\vspace{-1.5em}
      \begin{block}{Analyzing US TB Reduction Strategies}
        \begin{itemize}
          \item Implemented in \texttt{R}, with various numerical DE solvers
          \item Tracks US Health Care System (HCS) cost
          \item Tracks statistics about various health states
        \end{itemize}
      \end{block}

      \begin{block}{Intervention Analysis}
        %\vspace{-1.6em}
        \begin{figure}[h]
          \begin{minipage}[c]{0.6\textwidth}
            \includegraphics[height=0.7\textwidth,width=\textwidth]{redEnLTBIIncGrouped}
          \end{minipage}
          \hspace{0.5em}
          \begin{minipage}[c]{0.35\textwidth}
            \caption{Incidence/million in USB, FB, and total populations,
                     given 0\%, 50\%, 75\%, or 100\% treatment of incoming
                     LTBI}
          \end{minipage}
          \label{fig:redEnLTBI_incidence}
        \end{figure}
        \begin{figure}[h]
          \begin{minipage}[c]{0.6\textwidth}
            \includegraphics[height=0.7\textwidth,width=\textwidth]{incLTBItrmtIncGrouped}
          \end{minipage}
          \hspace{0.5em}
          \begin{minipage}[c]{0.35\textwidth}
            \caption{Incidence/million in USB, FB, and total populations,
                     given 0\%, 100\%, or 300\% LTBI treatment increase}
          \end{minipage}
          \label{fig:incLTBItrmt_incidence}
        \end{figure}
      \end{block}
      \begin{block}{Economic Modeling}
        \begin{itemize}
          \item Tracks treatment costs for various disease states
          \item Estimates implementation cost of intervention
          \item Active TB Costs: \$14,014.90 LTBI Costs: \$403.45

        \end{itemize}
        \vspace{.5em}
        %\begin{columns}[T]
          %\begin{column}{.66\textwidth}
            \begin{figure}[h]

              \begin{minipage}[c]{0.6\textwidth}
                \includegraphics[height=0.7\textwidth, width=\textwidth]{EnLTBIRedGroupCost.pdf}
              \end{minipage}
              \hspace{0.5em}
              \begin{minipage}[c]{0.35\textwidth}
                \caption{Cumulative implementation costs, US HCS savings, and net
                       US costs of LTBI arrival cure rates. Cost/case cured was
                       \$600, \$800, and \$1000 for 50\%, 75\%, and 100\%
                       cured}
              \end{minipage}
              \label{fig:redEnLTBI_costs} 
            \end{figure}
          %\end{column}
          %\begin{column}{.26\textwidth}
            % Base HCS Costs:
            % \begin{description}
            %   \item[Active TB:]\hfill \\ 
            %     \$14,014.90
            %   \item[LTBI:]\hfill \\ 
            %     \$403.45
            % \end{description}
          %\end{column}
        %\end{columns}
      \end{block}
    \end{column}

    \begin{column}{.3\textwidth}
      %\vspace{-.5em}
      \begin{block}{An Agent-Based Implementation}
        Agent-based models capture disease dynamics on the individual level and
        reflect stochasticity and granularity lost in compartmental models.
        Agent-based counterparts to the Hill model were implemented in Netlogo
        and \texttt{C++}.
        \begin{columns}
          %\vspace{-2em}
          \begin{column}{.45\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[width=\textwidth]{NLHMinc}
              \end{center}
              \caption{Incidence/million for R and NetLogo models (12 runs, $\Delta t$ = 0.1, popConst = 100)}
              \label{fig:NLHMinc}
            \end{figure}
          \end{column}
          \begin{column}{.45\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[width=\textwidth]{finalRunSmall2}
              \end{center}
              \caption{Incidence/million for R and C++ models (2100 runs, $\Delta t$ = 0.01, popConst = 1)}
              \label{fig:finalRun}
            \end{figure}
          \end{column}
        \end{columns}
      \end{block}

      \begin{block}{Stochastic Models as a Measure of Variability}
        \begin{columns}
          %\vspace{-2em}
          \begin{column}{.45\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[width=\textwidth]{IN0dist}
              \end{center}
              \caption{Distribution of USB Incidence (C++) with fitted Normal curve}
              \label{fig:IN0dist}
            \end{figure}
          \end{column}
          \begin{column}{.45\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[width=\textwidth]{IN1dist}
              \end{center}
              \caption{Distribution of FB Incidence (C++) with fitted Normal curve}
              \label{fig:IN1dist}
            \end{figure}
          \end{column}
        \end{columns}
      \end{block}

      \vspace{1.1em}
      \begin{block}{References}
        Hill, A. N., Becerra, J. E., \& Castro, K. G. (2012). Modelling tuberculosis trends in the USA. Epidemiology and infection, 140(10), 1862.
      \end{block}

    \end{column}
  \end{columns}
\end{frame}
\end{document}