### 2014-07-14 in_progress/presentations/paper/draft.tex
Working on draft. This paper is definitely getting out.

% arara: pdflatex
% arara: bibtex
% arara: pdflatex
% arara: pdflatex
\documentclass{amsart}
\usepackage[pdftex]{graphicx}
\usepackage{enumerate,placeins}
\usepackage{color}
\graphicspath{{figures/}}

\newcommand{\TODO}[1]{\hfill\\{\huge \color{red} TO DO:} #1 \hfill \\}

\title{Modeling Intervention Strategies for United States TB Control}
\begin{document}
\maketitle

{\huge \color{red} List of results}
Plan:
\begin{enumerate}
  \item Intervention Analysis: 
    \begin{enumerate}
      \item Reducing Entering LTBI
      \item Cutting Transmission
      \item Reducing Progression Rate
    \end{enumerate}
  \item Qualitative Strategy Conclusions:
    \begin{enumerate}
      \item Sourced Incidence \& Cost
      \item Plateauing Effect
      \item Extremal values for entering LTBI reduction
      \item Linearity in Cost/Incidence for entering LTBI reduction
      \item CPCA data
    \end{enumerate}
  \item Statistical Verification/Agent Based Feasibility
    \begin{enumerate}
      \item Normal Distributions
      \item Basic Reproductive Numbers
      \item Agent based running time \& Extendibility. 
    \end{enumerate}
\end{enumerate}
Ideas:
\begin{itemize}
  \item Plateauing Effect
  \item Intervention Analysis
  \item Statistical Analysis of Hill Model
  \item Agent-Based Feasibility
  \item Theoretical/Experimental justification of basic reproduction number
  \item Overall qualitative strategy conclusions. 
\end{itemize}

\section{Abstract}
An epidemiological and economic model for tuberculosis in the United States
 was developed, extending a model developed in 2012 by the Centers
for Disease Control and Prevention. Sensitivity analysis of this model reveals
that the percentage of foreign born immigrants with LTBI is a highly influential
parameter. Various intervention strategies, particularly those that reduced LTBI
among foreign-born arrivals, were evaluated, and cost per case averted estimates
were found.  Additionally, a population-level, stochastic, agent-based model was
built. This model demonstrated the feasibility of using agent-based modeling in
the context of minimal population heterogeneity and provided statistical
characterization of the results of the compartmental model. The distributions
of the final population sizes and incidence rates projected by the stochastic model
were found to be approximately normal, with means centered at the results of
compartmental model.

\section{Introduction}
Epidemiological models allow public health professionals to predict and analyze
disease dynamics and intervention effectiveness. The most common examples of
such models are detereministic, compartmental, differential equation models. In
these models, the population is split between several possible health states, or
compartments, and a system of differential equations governs population flow
between the various compartments. In 2012, Hill, Becerra, and Castro implemented
a compartmental differential equaiton model of tuberculosis (TB) in the United
States (US).  Their model characterized five health states across two
subpopulations, US-born (USB) and foreign-born (FB), for a total of 10
compartments. They used this model to evaluate several possible intervention
strategies and ultimately concluded that though increasing LTBI treatment was an
effective intervention strategy, the US was unlikely to meet their stated goal
of elimination of TB in the US by 2100. Here, we present an extended Hill model
with additional tracking capabilities, such that it can now report
further granularity in the disease dynamics. Furthermore, this model also tracks
economic data in order to project the US health care system (HCS) costs due to
TB given our current policy, as well as in the context of various interventions.

{\huge \color{red} Fix this paragraph. Its not the best. \textbackslash/}

Though compartmental, differential equation models are the most common strategy
when modeling large-scale diseases at the population level, an alternative
approach is to use an agent-based model. Agent
based models provide a more nuanced view of disease spread as they provide a
more biologically accurate modeling framework; however, they are often
considered to be computationally infeasible at population level. In order to
challenge this assumption of infeasibility and provide statistical verification
of the results of the compartmental model, we implemented a stochastic,
agent-based model of US TB dynamics at the population level.

\section{Background}
\subsection{Tuberculosis} x 
\TODO{Use TB instead of Tuberculosis maybe?}
Tuberculosis is an infectious bacterial disease that primarily affects the lungs,
with common symptoms including severe coughing, fevers, and chest pain.
Tuberculosis is a major global health concern; The World Health Organization
estimates that 8.6 million developed TB and 1.3 million people died from
the disease in 2012. Tuberculosis is an airborne disease and is typically
transmitted from person-to-person by coughing, sneezing, and the exchange of
other respiratory fluids.  Infected individuals first enter a latent phase,
during which the disease causing agent, \emph{Mycobacterium tuberculosis}, is
present in their bodies, but the individuals remain asymptomatic and
noncontagious. During this phase, individuals are said to have latent TB
infection (LTBI). Eventually, some number of individuals with LTBI progress into
what is known as the active phase of the disease, or active TB. Individuals with
an active TB infection are symptomatic and contagious. The lifetime risk of
progressing from Latent TB Infection (LTBI) to active TB is approximately
5-10\%, although individuals with LTBI have the greatest chance of developing
Active TB during the first two years immediately following infection, and if
they have a suppressed immune system, such as in the case of a comorbid HIV
infection. Cases of Active TB may be classified as pulmonary ($\sim$80\% of
cases) or non-pulmonary ($\sim$20\% of cases), depending upon the biological
location of infection. In the Hill model, pulmonary Active TB is treated as
contagious and non-pulmonary Active TB is treated as non-contagious, which is an
assumption we also use here.

In the United States, 9,951 new cases of TB were reported in 2012 by the Centers
for Disease Control. The disease burden varied for different subpopulations in
the U.S., and, in particular, the incidence rate of TB was 11.5 times higher for
the FB population than the USB population.  For the
past 20 years, the overall incidence of TB in the U.S. has declined each year,
but in 2012 the annual incidence level of 32 new cases/million remains above the
desired goal of the $<$1 new case/million, which is the target for disease
elimination set by the Centers for Disease Control.

\TODO{}citations: CDC Tuberuclosis 2013 Report - http://www.cdc.gov/mmwr/preview/mmwrhtml/mm6211a2.htm
WHO Global Tuberculosis Report 2013 - http://www.who.int/tb/publications/global\_report/en/

\subsection{The Hill Model}
A flowchart representation of the Hill Model is shown in
Figure~\ref{fig:hillModelSchematic}. Each compartment represents a different
possible TB health state for every US-born or FB individual, and
arrows between different compartments represent possible transitions between
states. Individuals also leave the model from all compartments due to natural
death, which is left out of the figure for clarity.

\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=0.25]{figures/HillModelFlowChart.png}
  \end{center}
  \caption{Schematic of the Hill Model.}
  \label{fig:hillModelSchematic}
\end{figure}

The majority of USB and FB individuals fall into the Susceptible (S) category,
which includes everyone who is uninfected and has not been exposed to TB, or
those who have been successfully cured of the disease. After exposure to an
individual with TB, a person in the Susceptible compartment can develop Latent
TB Infection (LTBI). Latently infected individuals are asymptomatic and
non-infectious, but have some risk of developing active TB infection over time.
To reflect the fact that real LTBI patients have a much higher risk of
developing active TB within two years of exposure, the Hill Model splits the
LTBI compartment into Acute LTBI (fast progressors) and Chronic LTBI (slow
progressors). Accordingly, individuals in the Acute LTBI compartment have a
higher risk of developing active TB than those in the Chronic LTBI compartment.
Individuals in the Chronic LTBI compartment may also be exogenously re-infected
and transition to the Acute LTBI compartment.

Latently infected individuals may progress to one of two active TB states: Infectious TB (I) or 
Non-Infectious TB (J).  Individuals in both compartments have an increased risk of death from
active TB infection, but only individuals in the Infectious TB compartment are contagious.  
In addition, individuals in all of the infected compartments (F, L, I, J) may be treated or self-cure
themselves of their respective TB health condition.  However, in the model, treatment or self-cure from TB
does not grant immunity, and all healthy individuals are grouped in the Susceptible compartment
and may be re-infected at a later time.\\

\section{Methods}
\subsection{Basic Structure}
Both the Hill Model (Basic Hill Model) and our extended Model (Extended Hill
Model) were implemented in \texttt{R} as a system of differential equations,
which were solved via the $\texttt{lsoda}$ routine. The systems of differential
equations and relevant constants used to capture basic disease dynamics in the
Basic Hill Model are shown in Figures~\ref{fig:hillEquations}
and~\ref{fig:basicHillConstants}, though a full analysis of the model is better
suited for \TODO{<+Cite Hill+>}. \TODO{Add equations for $\lambda$. I don't have
a pdf of Hill, apparently, though, so I can't.}  In the Extended Hill Model,
each flow rate equation from the Basic Hill Model was separated into the
components that tracked specific population migrations (i.e., terms that dictate
flow rates from one compartment to another) and these were tracked separately.
In other words, rather than tracking net flow into various compartments, the
Extended Hill Model tracks flow between all pairs of compartments separately.
Furthermore, these equations were each also supplemented to estimate their
corresponding US HCS costs. A relevant subset of the Extended Hill equations are
detailed in the appendix, section~\ref{sec:extendedHillEqs}. Historical data
from 2000-2008 was used to initialize the model, which was then run up to the
year 2100.

\begin{figure}
  \begin{align*}
    \dot{S}_0 &= \rho(N_0+N_1) + \sigma_0^FF_0 + \sigma^LL_0 
                +\varphi_0(I_0+J_0) - \lambda_0 S_0 - \mu_0 S_0\\
    \dot{F}_0 &= p\lambda_0S_0 + xp\lambda_0L_0 - (\mu_0+\nu^F+\sigma_0^F)F_0\\
    \dot{L}_0 &= (1-p)\lambda_0S_0 - xp\lambda_0L_0 
                -(\mu_0+\nu_0^L+\sigma^L)L_0\\
    \dot{I}_0 &= q(\nu^FF_0 + \nu_0^LL_0) - (\mu_0+\mu^d+\varphi_0)I_0\\
    \dot{J}_0 &= (1-q)(\nu^FF_0 + \nu_0^LL_0) - (\mu_0+\mu^d+\varphi_0)J_0\\
    \dot{S}_1 &= (1-f)\alpha(N_0+N_1) + \sigma_1^FF_1 + \sigma^LL_1 
                +\varphi_1(I_1+J_1) - \lambda_1S_1 - \mu_1S_1\\
    \dot{F}_1 &= gpf\alpha(N_0+N_1) + p\lambda_1S_1 + xp\lambda_1L_1 
                -(\mu_1+\nu^F+\sigma_1^F)F_1\\
    \dot{L}_1 &= (1-gp)f\alpha(N_0+N_1) + (1-p)\lambda_1S_1 - xp\lambda_1L_1 
                -(\mu_1+\nu_1^L+\sigma^L)L_1\\
    \dot{I}_1 &= q(\nu^FF_1 + \nu_1^LL_1) - (\mu_1+\mu^d+\varphi_1)I_1\\
    \dot{J}_1 &= (1-q)(\nu^FF_1 + \nu_1^LL_1) - (\mu_1+\mu^d+\varphi_1)J_1\\
  \end{align*}
  \caption{System of Differential Equations given in the Hill Model.}
  \label{fig:hillEquations}
\end{figure}
\begin{figure}
  \begin{align*}
    \sigma^{L} &= 0.057   &&\text{(Treatment rate for chronic LTBI)}\\
    v^{L}_{0}  &= 0.0014  &&\text{(Progression rate for reactivation in the USB
                                   population)}\\
    v^{L}_{1}  &= 0.0010  &&\text{(Progression rate for reactivation in the FB
                                   population)}\\
    f          &= 0.187   &&\text{(Fraction of FB arrivals with LTBI)}\\
    p          &= 0.103   &&\text{(Fraction of new infections which are acute)}\\
    ARI_{0}    &= 0.00030 &&\text{(2000 Annual Risk of Infection, USB)}\\
    q          &= 0.708   &&\text{(Fraction of infections progressing to infectious
                                   disease)}\\
    g          &= 0.0047  &&\text{(Fraction of FB arrivals with LTBI who are fast
                                   progressors)}\\
    \sigma^{F} &= 0.461   &&\text{(Cumulative fraction of treatment for acute
                                   infection)}\\
    r_{0}      &= 0.667   &&\text{(Fraction of cases due to reactivation in USB
                                   population)}\\
    r_{1}      &= 0.780   &&\text{(Fraction of cases due to reactivation in FB
                                   population)}\\
    \mu^{d}    &= 0.115   &&\text{(Mortality rate due to TB)}\\
    x          &= 0.111   &&\text{(Fraction of re-infected chronic LTBI moving to
                                   acute infection)}\\
    \phi       &= 0.897   &&\text{(Cumulative fraction self-cure and treatment of
                                   active disease)}\\
    e_{0}      &= 0.965   &&\text{(Fraction of preferred contacts within own
                                   population for USB)}\\
    e_{1}      &= 0.985   &&\text{(Fraction of preferred contacts within own
                                  population for FB)}\\
  \end{align*}
  \caption{The relevant constants used in the Basic Hill Model.  }
  \label{fig:basicHillConstants}
\end{figure}

The vector variables $S_0, F_0, L_0, I_0, J_0$ contain the number of US-born
individuals in the S, F, L, I, J compartments respectively, whereas $S_1, F_1,
L_1, I_1, J_1$ contain the number of foreign-born individuals. $N_0$ and $N_1$
are the total populations of US-born and foreign-born individuals. 
\TODO{Is the next bit needed? I think move the stuff below out and the stuff
above back where we introduce the equations.}
The constants $\rho$ and $\alpha$ are birth rates, while $\mu_i$, and $\mu_d$
are death rates. A complete list and descriptions of all constants used in the
model can be found in the Appendix, Section~\ref{subsec:hillConstants}.

\subsection{Additional Tracking Capabilities}
In order to refine the tracking capabilities of the Hill Model, the original
differential equations used to describe TB spread were separated into their
component parts and each part was tracked separately. In effect, each flow rate
between two compartments in the Basic Hill model was made into its own
``virtua'' compartment, thereby enabling computation not only of expected TB
incidence, but also of the source of that incidence. In particular, progressions
into active TB due to activations of LTBI or exogenous infection were tracked,
allowing for sourced incidence data to be generated. These equations allowed the
model to track the sourced number of TB cases, TB deaths, and natural deaths.
Furthermore, the model also tracks the sourced cost on the US HCS. The
immigrating cases of LTBI (acute or chronic) were also tracked. In the case of
intervention testing, the number of cured and untreated cases of entering LTBI
were both tracked. Cured cases of LTBI entering, TB deaths, total TB cases, and
total cost were also tracked discounted at 3\% annually. This discounting was
converted to a continuous differential equation for use within the model. In
general, incidence data is calculated in the same way in the Extended Hill Model
as it was in the Basic Hill Model.

\TODO{Where does this fit in? I'm not seeing why this is here? Maybe in the
results, instead?}
In addition, estimations of the basic reproductive number of FB or USB active,
infectious TB cases were made from a theoretical and an experimental
perspective. Experimental data were calculated by reducing the initial
FB or USB infectious TB populations by one and allowing the
model to progress otherwise as normal. The decreased number of total TB cases
seen details how many infections can be thought to be due to one infected
individual in the given population. In the agent-based model, this \emph{one}
was a true model agent whereas in the compartmental model, this decrease of
\emph{one} was a proportional decrease based on the scaled population level.

From a theoretical perspective, the spread of TB was thought to be a geometric
series. If one infectious individual infects $x$ people annually, over the
course of $N$ years the total number of infections caused by this individual can
be approximated by the geometric series of $N$ terms with rate $x$ and initial
term $1$. In this case, $N = 100$, and the ratios for FB infectious individuals
and USB infectious individuals were obtained from 
\TODO{<+RATIO CITATION COLIN+>.}
\subsection{Economic Modeling}
In order to track the US HCS TB cost, we made the following treatment cost
assumptions. A single case of active TB was assumed to yield a \$14,000 US HCS
cost, charged immediately upon disease contraction. This cost is the weighted
mean of the costs of cases requiring hospitalization (49\%) and cases not
requiring hospitalization (51\%).  These cost estimates were found in 
\TODO{<+DYLAN COST STUDY CITATION+>.} LTBI treatment costs were assumed to be
\$468.00, charged immediately upon successful treatment. These estimates were
calculated based on the cost of a successful treatment, and the typical
adherence and efficacy of said treatment 
\TODO{<+DYLAN COST STUDY CITATION, ADHERENCE CITATION+>.}
It is important to note that the charging assumptions are different for
treatment of active TB as opposed to treatment of latent TB. In particular,
cases of active TB imply a cost immediately upon disease contraction, whereas
latent TB costs are charged only upon successful treatment. The motivation for
this distinction is that active TB treatment is mandated by the US for all known
cases and, further, most treatment costs are incurred at the beginning of the
treatment cycle. As such, charging for every known case, immediately at
contraction is a well motivated assumption. On the other hand, latent TB is only
rarely treated in discovered cases and the costs are more evenly distributed.
Furthermore, the choice to charge only upon successful treatment yields a more
conservative estimate of the success of any intervention analyzed. Finally, it
is worth noting that, though accurate, these estimates of US HCS treatment cost
are not intended to be fixed parameters. The code for this model is freely
available, and as new research amends the necessary costs and new methods of
treatment are developed, these values can be modified and the model re-run to
produce a revised prediction of the US HCS TB cost.

In addition to refining the model to produce sourced US TB incidence estimates,
the additional cost modeling was also tracked in a way that preserved the source
of the US HCS cost. In particular, active TB treatment cost was tracked
separately for cases resulting from an activation of LTBI (modeled as a
progression from the Chronic Latent category) and for cases resulting from a
novel, exogenous infection (modeled as a progression from the Acute Latent
category). All cost equations were also solved by $\texttt{lsoda}$.  Given the
uncertainty in estimates of LTBI treatment cost, extensive sensitivity analysis
was performed on this parameter relative to final cost outcomes. Furthermore,
the US HCS cost caused by LTBI treatment (which was tracked) was often omitted
from Cost Benefit intervention analyses, so as to provide the most conservative
estimate of total US HCS cost given various TB interventions. In addition to US
HCS cost, the system was implemented so as to also track projected intervention
implementation cost, given user-inputted parameters relating to various possible
intervention cost strategies. The extended Hill Model was used to track the
effect of many interventions tested by the Hill Model. In particular, the
intervention strategy of curing entering cases of LTBI proved very promising and
elimination year, final cost, and cost per case averted were tracked for various
levels of entering LTBI cure rate. The interventions were implemented so as to
take effect during the year 2013 and run to 2100. The numerical DE solver
$\texttt{lsoda}$ used was run with a time step of $0.8$. Sensitivity analysis
was performed on this parameter and reducing the time step was shown to have
minimal effect on final size or cost values.

\subsection{Agent Based}
\TODO{What's the story we are trying to tell here? I mean, we are following a
  very standard format, but I think it would be better if we had a cohesive
  message.}
The population level agent based model was implemented in several ways. Early
implementations were built in \texttt{Netlogo} and \texttt{Java}, but final
implementations were constructed in \texttt{C++}. Probabilities of agent
progression between various health states were computed given rates in the hill
model and a variety of integration approximations. The model maintained
individual records for every individual with LTBI or active TB. Susceptible
individuals were not modeled as agents; instead, a binomial distribution was
used to probabilistically determine the number of new infections each time step.
Similarly, the number of new immigrants each time step was chosen from a
probability distribution. As the model was stochastic, all final data is
inherently merely a probabilistic sample; as such, many simulations are required
to draw conclusions from the model. As such, we collected final population
health statistics from approximately 2160 model simulations, with each agent in
the model truly representing one infected individual, and a time step of $0.01$.
These data were analyzed in $R$. The agent based model also tracks deaths, total
TB cases, and infection sources, and can also easily track economic data. These
data were consistently normal and matched the deterministic model. 
\TODO{Does it track economic data yet?}

Further implementations of the \texttt{C++} model were made that deviated from
the basic Hill Model. Specifically, the acute latent compartment in the basic
Hill Model is a vestigial necessity of compartmental modeling and not reflective
of the biology of tuberculosis. As such, an agent based model was implemented
that more closely respects the biology of tuberculosis. Results from this model
were also normal and did not differ significantly from the strict Hill model.

\section{Results}
\subsection{Basic Population Breakdown}
The additional tracking capabilities offered several key insights into US TB
dynamics. Figure~\ref{fig:incPlotSourced} shows the yearly incidence of US TB, broken
down into infection source. It can be seen that the majority of the US TB load
is driven by activations of FB LTBI, followed by USB LTBI activations. This data
further agrees with the conclusions drawn by Hill, Becerra, and Castro about the
necessity of LTBI treatment in any effective intervention strategy. 
\TODO{Why does it agree? We aren't doing any data analysis here, which is a
problem. We need to walk our readers through this.}
\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=.5]{incPlotSourced}
  \end{center}
  \caption{Sourced yearly incidence data generated by the extended Hill Model.}
  \label{fig:incPlotSourced}
\end{figure}
Further, figure~\ref{fig:costPlotSourced} shows a similarly sourced plot, but analyzing the
final US HCS costs due to active TB. One can see that in this plot, roughly half
of the US TB HCS costs are due to activations of LTBI. Note that in this data,
it is appropriate to think of the costs due to acute LTBI as costs due to
exogenous re-inection, whereas costs due to chronic LTBI are activations of long
standing latent infections. 
\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=.35]{costPlotSourced}
  \end{center}
  \caption{Sourced US HCS economic TB load. Note that this data only illustrates
    the load due to treating active TB, but illustrates where the infections
    driving this cost come from.}
  \label{fig:costPlotSourced}
\end{figure}
Note that both of these plots underestimate the impact of LTBI on the spread of
TB, as every LTBI activation contributes not only
to incidence and costs directly, but also indirectly by causing additional
future cases, which is not captured in these graphs. Further, there are also US
HCS costs due to LTBI treatment, which is not illustrated in these graphs. 
\subsubsection{Basic Reproduction Number}
The basic reproduction number of FB or USB cases of infectious TB was also
estimated by this system. Using the best-fit values of parameters from the 
Hill model, we found the average number of secondary infections arising from
a single case of infectious TB to be 0.423 and 0.370 for USB and FB populations
respectively.  These values differ because in the model, there are different
estimates for death rates and the rate of TB treatment and self-cure in the USB and FB populations, 
so the average infectious period differs between the two populations.
Extrapolating these results, we can think of the
total number of secondary infections over 100 years due to a FB or USB
infectious TB case as describing a geometric series in a large population.
Presuming there are no overlaps in infectious contacts, if a single case of
infectious TB in either population infects $p_f, p_u$ new cases,
respectively, then we can say that over the course of 100 years, the total
number of cases infected will follow a geometric series.
{\huge \color{red} TODO: formalize this with math. We can fit it here.}
This analysis predicts
that over 100 years, one USB infection will lead to $1.04$ subsequent infections,
whereas one FB infection will lead to $.83$ subsequent infections. Experimentally, 
these data were also analyzed, with results of $1.03$ and $.64$, respectively. 
Full calculations are included in the Appendix.

\subsection{Intervention Analysis}
The primary interventions analyzed by the extended Hill were those that analyzed
curing various percentages of entering LTBI cases. Four indicative percentages
chosen were $5\%$, $10\%$, $25\%$, and $50\%$. Note that the Hill model does not
distinguish between documented and undocumented immigration, so these
percentages are over \emph{all} immigration into the United States, and can thus
be thought to represent significantly higher percentages of documented
immigration. It was seen that in this model as well as in the Basic Hill Model,
no analyzed intervention predicted elimination by 2100. In order to obtain
elimination by 2100, at least 95\% of entering LTBI cases had to be cured, which
is infeasible in the real world. However, it was seen that curing entering cases of
LTBI resulted in a net US HCS cost per case averted of \$67,654.07 at 2025, \$28,699.15 at 2050,
and \$17,912.95 at 2100 (assuming 25\% reduction with each cure costing \$800; see Table
\ref{tab:cpcaArt} for other percentages). Given the variable nature of LTBI
treatment cost, the model code is extendible such that a user can adjust these
costs themselves to explore more specific methods of curing entering LTBI,
especially as the treatment methods and associated costs evolve in future years.
Further, it was also found that the relationship between total incidence at 2100
and percentage of incoming LTBI cases cured was linear, and from this estimates
were made of the yearly average US HCS savings garnered by curing one case of
entering LTBI over the time scale 2000 to 2025, 2000 to 2050, and 2000 to 2100.
This value peaked at \$1.283 billion at 2100 (25\% reduction). This illustrates
that it would be cost saving to cure cases of LTBI at the cost of \$1.283
billion ``2000'' dollars over the time period 2000-2100. These intervention
strategies also resulted in 11,900; 29,880; and 60,189 fewer cases of TB seen in
the US, and 1,025; 2,573; and 5,185 fewer TB deaths, for 10\%, 25\%, and 50\%
reduction, respectively.
{\huge \color{red} TODO: This section is a bit confusing. Also, a graph would
make things much prettier.}

% latex table generated in R 3.0.1 by xtable 1.7-1 package
% Mon Oct 14 00:36:15 2013
\begin{table}
\centering
\begin{tabular}{|r|cccc|} \hline
       & red5     & red10    & red25    & red50    \\ \hline
  %2000 & ND       & ND       & ND       & ND       \\ 
  2025 & 67908.81 & 67855.46 & 67654.07 & 67332.53 \\ 
  2050 & 28938.23 & 28878.70 & 28699.15 & 28404.13 \\ 
  2075 & 21044.88 & 20989.16 & 20820.53 & 20541.77 \\ 
  2100 & 18130.11 & 18076.11 & 17912.95 & 17642.64 \\ \hline
\end{tabular}
\caption{Cost Per Case Averted by Reducing Incoming LTBI by X percent (in dollar per case)} 
\label{tab:cpcaArt}
\end{table}

Several other intervention strategies analyzed by the Hill Model were also
examined via the Extended Hill, and economic properties about each were tracked.
The results for these interventions, which were consistently less effective than
curing entering LTBI cases, are given in the appendix. {\huge \color{red} TODO:
Which section?}

\subsection{Sensitivity Analysis}
To account for uncertainty of input parameter values and to gain insight about
the most influential parameters in the model, we used Latin Hypercube Sampling
to analyze the deterministic model, implemented with the \texttt{lhs} package in
\texttt{R}.  We varied the 16 parameters from the original Hill model according
to triangular distributions centered around their best-fit values, and two
additional treatment cost parameters according to uniform distributions.
{\huge\color{red} TODO: These distributions were chosen because\ldots}. All
non-economic parameters analyzed reported statistically similar results to their
analysis in the Basic Hill Model. For economic outcomes, we found the parameter
$f$, the fraction of FB arrivals with LTBI, to be highly correlated with the
projected overall cost burden of TB in the United States over the next 100
years.  Another highly influential parameter in the model was $\sigma^{L}$, the
treatment rate for chronic LTBI.  However, while increasing $\sigma^{L}$
significantly decreases the cost burden due to Active TB treatment, it increases
the cost burden due to Latent TB treatment by a greater amount, given the
treatment cost estimates of \$700 and \$14,000 for LTBI and Active TB,
respectively.  A full description of the Latin Hypercube Sampling analysis and
tables of Partial Rank Correlation Coefficient (PRCC) results for all parameters
are given in the appendix, in Section~\ref{subsec:lhs}.

\subsection{Agent Based Evaluation}
The agent-based model allowed the statistical properties of the system to be
analyzed and verified. In particular, it illustrated that the deterministic
Hill model provides a robust and consistent statistical measure of TB epidemic
behavior in the US conditions. We found that the distribution of incidence and
final population sizes were normal, with mean accurate to the deterministic
model and standard deviations given in table {\huge\color{red} TODO: Add table
\# and section \# here}, in the Appendix. Additionally,
the agent-based model provided a computational framework to produce
meaningful intervention results. On student lab gradet hardware, a statistically
meaningful experiment could be run overnight, producing data in one day. This
demonstrates that this modeling strategy is feasible in this general case.
Furthermore, the agent-based model provides a greater extensibility to more
specific and biologically reasonable scenarios, such as geographic dependency or
greater population heterogenity.

\section{Discussion}
These results confirm the hypothesis that curing incoming LTBI rates is a
necessary step towards elimination and indicate that it is a cost effective
option. \\
{\huge \color{red} TODO: Where did we show it was cost effective relative to other
interventions? Also, the section below (future work) might be better just
directly included in the discussion. Also, if we want to make this claim here we
need to include other interventions in the main body.}\\
In addition, beyond demonstrating quantitatively that this intervention strategy
is a necessary step towards US elimination, we also obtained qualitative
indications that this is the case. In every intervention we tested (see, for
example, Intervention \TODO{<+ADD ME+> and <+ADD ME+> in section <+ADD ME+>}) we
observed a plateauing effect, where the incidence per million stabilized after
some time and resisted further change. In particular, even when transmission was
reduced to $0\%$, we still observed this plateauing effect, which implies that
something other than active transmission was generating enough incidence to
sustain TB in an endemic state in the US population. However, when the FB LTBI
immigration was cut, this effect vanished. These results therefore support the
argument that the USB LTBI population is sufficiently large to sustain a
sizable, endemic TB-disease state in the United States provided it is
continually refueled via the entering LTBI immigrants. Thus, there is no
intervention that can ignore the LTBI population that can hope to eliminate TB
in the United States. Among those interventions that treat cases of LTBI,
targeting the immigrating LTBI requires no additional US HCS dollars beyond what
is spent now to identify the LTBI population and is therefore extremely
efficient. Additionally, our analysis shows that paying for the treatment of
these individuals is extremely cost effective {\huge \color{red} TODO: Show
this}.
\subsection{Future Work}
This work could be extended by examining different classes of interventions or
more accurately estimating intervention cost with the deterministic extended
Hill. Further work could also be done with the agent-based Hill Model, by using
it to examine the effect contact structure plays on US TB incidence levels or to
examine the effects drug-resistant TB will have on US TB dynamics.

\section{Appendix} 
\subsection{Hill Constants}
\label{subsec:hillConstants}
Below, we detail some of the relevant constants in the Basic Hill Model, and
their best-fit values. A full listing of constants used in the original Hill
Model can be found in \TODO{<+CITATION HERE+>} \TODO{Can we delete this
section?}

\textit{(USB = US born, FB = Foriegn born)}
\begin{align*}
  \sigma^{L} &= 0.057   &&\text{(Treatment rate for chronic LTBI)}\\
  v^{L}_{0}  &= 0.0014  &&\text{(Progression rate for reactivation in the USB
                                 population)}\\
  v^{L}_{1}  &= 0.0010  &&\text{(Progression rate for reactivation in the FB
                                 population)}\\
  f          &= 0.187   &&\text{(Fraction of FB arrivals with LTBI)}\\
  p          &= 0.103   &&\text{(Fraction of new infections which are acute)}\\
  ARI_{0}    &= 0.00030 &&\text{(2000 Annual Risk of Infection, USB)}\\
  q          &= 0.708   &&\text{(Fraction of infections progressing to infectious
                                 disease)}\\
  g          &= 0.0047  &&\text{(Fraction of FB arrivals with LTBI who are fast
                                 progressors)}\\
  \sigma^{F} &= 0.461   &&\text{(Cumulative fraction of treatment for acute
                                 infection)}\\
  r_{0}      &= 0.667   &&\text{(Fraction of cases due to reactivation in USB
                                 population)}\\
  r_{1}      &= 0.780   &&\text{(Fraction of cases due to reactivation in FB
                                 population)}\\
  \mu^{d}    &= 0.115   &&\text{(Mortality rate due to TB)}\\
  x          &= 0.111   &&\text{(Fraction of re-infected chronic LTBI moving to
                                 acute infection)}\\
  \phi       &= 0.897   &&\text{(Cumulative fraction self-cure and treatment of
                                 active disease)}\\
  e_{0}      &= 0.965   &&\text{(Fraction of preferred contacts within own
                                 population for USB)}\\
  e_{1}      &= 0.985   &&\text{(Fraction of preferred contacts within own
                                population for FB)}\\
\end{align*}
\subsection{The Extended Hill Equations}
\label{sec:extendedHillEqs}
Below, we detail some of the relevant equations used in the extended hill model.
In many cases, the equations of the Hill model were split into their component
parts, which could then be tracked separately.

\TODO{What equations should be put in here?}

Here, a continuous discounting approximation is used to a discount rate of $3\%$
yearly, found via $discV$. In many of the differential equations, a small $c$ is
used to denote \emph{cost of}. 
\FloatBarrier
\subsection{Estimations for Infectious Rates of FB and USB}

\subsection{Statistical Qualities of the Agent Based Model}

\begin{figure}
    \includegraphics[scale=0.4]{figures/qqnormUSBInc.pdf}
    \includegraphics[scale=0.4]{figures/qqnormFBInc.pdf}
    \includegraphics[scale=0.4]{figures/qqnormUSBPop.pdf}
    \includegraphics[scale=0.4]{figures/qqnormFBPop.pdf}
  \caption{Normal Q-Q Plots for 2103 runs of the Stochastic Model with population constant = 1.}
  \label{fig:qqnormPlots}
\end{figure}

Results for USB Incidence, FB Incidence, USB Population, and FB Population in the year
2100 shown, overlayed with linear (normal) fit are shown in
figure~\ref{fig:qqnormPlots}. These plots demonstrate the normality of the
Stochastic data. 
\subsection{Finer Intervention Analysis}

% latex table generated in R 3.0.1 by xtable 1.7-1 package
% Mon Oct 14 00:36:15 2013
\begin{table}
\centering
\begin{tabular}{|r|cccc|} \hline
       & red5    & red10    & red25    & red50    \\ \hline
  %2000 & 0.00    & 0.00     & 0.00     & 0.00     \\ 
  2025 & 1115.12 & 2231.70  & 5592.43  & 11227.57 \\ 
  2050 & 3406.41 & 6821.27  & 17117.18 & 34447.71 \\ 
  2075 & 4976.77 & 9967.15  & 25021.33 & 50389.58 \\ 
  2100 & 5942.02 & 11900.90 & 29880.03 & 60189.49 \\ 
   \hline
\end{tabular}
\caption{Cases of TB Averted by Reducing Incoming LTBI by various percentages
         (Cure=\$800)} 
\label{tab:caRed}
\end{table}

% latex table generated in R 3.0.1 by xtable 1.7-1 package
% Mon Oct 14 00:36:15 2013
\begin{table}
\centering
\begin{tabular}{|r|cccc|} \hline
       & red5     & red10    & red25    & red50    \\ \hline
  %2000 & ND       & ND       & ND       & ND       \\ 
  2025 & 67908.81 & 67855.46 & 67654.07 & 67332.53 \\ 
  2050 & 28938.23 & 28878.70 & 28699.15 & 28404.13 \\ 
  2075 & 21044.88 & 20989.16 & 20820.53 & 20541.77 \\ 
  2100 & 18130.11 & 18076.11 & 17912.95 & 17642.64 \\ \hline
\end{tabular}
\caption{Cost Per Case Averted by Reducing Incoming LTBI by various percent (in
dollars per case, Cure=\$800)} 
\label{tab:cpcaRed}
\end{table}

\subsection{Latin H ypercube Sampling}
\label{subsec:lhs}
%The model parameters and variable names referred to in this analysis are listed
%below, along with best-fit or estimated values. 
Following the example of the
Hill model, we generated a Latin Hypercube Sample varying 18 of the input
parameters.  From these parameters, 16 are identical to the input parameters
varied in the sensitivity analysis of the original Hill model, and the remaining
two parameters, $C_{A}$ and $C_{L}$, detailed below, are variables for the average cost of
Active and Latent TB treatment in the US.   Probability distributions for the
original 16 parameters of the Hill model were all set to be Triangular, with
mode at the best fit value and end points at the 2.5 and 97.5 percentile values
reported in the Hill model.  Probability distributions for $C_{A}$ and $C_{L}$
were set to be Uniform, with range +/- 10\% of the estimated value.  All
probabilty distributions used to generate the Latin Hypercube are listed in
Table 1.\\

{\bf Additional Cost Parameters}
\begin{align*}
  C_{A} &= \$14,014.50 &&\text{(Cost per Active TB treatment)}\\
  C_{L} &= \$700       &&\text{(Cost per Latent TB treatment)}
\end{align*}
\vspace{5mm}

\begin{table}[h]
\centering
\begin{tabular}{l l}
\hline\hline\\
Parameter & Distribution\\ [0.5ex]
\hline\\
$\sigma_{L}$  & Tri(0.015,0.057,0.086) \\
$v^{L}_{1}$   & Tri(0.0009,0.0010,0.0014) \\
$f$                 & Tri(0.157,0.187,0.232) \\
$p$                & Tri(0.053,0.103,0.137) \\
$ARI_{0}$      & Tri(0.00021,0.00030,0.00030) \\
$q$                & Tri(0.569,0.708,0.825) \\
$g$                & Tri(0.0008,0.0047,0.0815)  \\
$\sigma_{F}$ & Tri(0.419,0.461,0.574) \\
$r_{1}$          & Tri(0.759,0.780,0.831) \\
$r_{0}$          & Tri(0.623,0.667,0.694) \\
$\mu^{d}$      & Tri(0.071,0.115,0.231) \\
$x$                 & Tri(0.088,0.111,0.860) \\
$v^{L}_{0}$   & Tri(0.0011,0.0014,0.0015) \\
$\phi$            & Tri(0.861,0.897,0.938) \\
$e_{0}$          & Tri(0.853,0.965,0.995) \\
$e_{1}$          & Tri(0.877,0.985,0.999) \\
$C_{A}$           & Uniform(12613,15416) \\
$C_{L}$           & Uniform(630,770) \\ [1ex]
\hline
\end{tabular}\\[1ex]

{\bf Table 1.} Probability distributions for model parameters, where Tri(x,y,z) denotes the 
Triangular distribution with endpoints (x,z) and mode y.
\end{table}

With a random Latin Hypercube Sample of size n=100,000, we computed partial rank correlation coefficients (PRCC) for each of the initial parameters and treatment costs, according to four different outcomes: 1) projected annual incidence in 2100 in the overall population, 2) projected cumulative cost of Latent TB treatments by 2100, 3) projected cumulative cost of Active TB treatments by 2100, 4) projected cumulative total cost of TB treatments by 2100.  For outcome 1, PRCC values are shown alongside PRCC values computed in the original Hill model for the same outcome in Table 2, showing the closeness of our findings to the sensitivity results of the original Hill model.  PRCC values for the remaining outcomes are reported in Table 3.  \\

\begin{table}[h]
\centering
\begin{tabular}{l r r}
\hline\hline\\
Parameter & Extended Hill Model & Original Hill model \\ [0.5ex]
\hline\\
$\sigma^{L}$  & -0.9303 & -0.9381 \\
$v^{L}_{1}$   & 0.7871  & 0.8309 \\
$f$                 & 0.7050  & 0.8072 \\
$p$                & 0.8369  & 0.6100 \\
$ARI_{0}$      & 0.5950  & 0.4939 \\
$q$                & 0.5797  & 0.4543 \\
$g$                & 0.6122  & 0.4517 \\
$\sigma^{F}$ & -0.4911 & -0.3772 \\
$r_{1}$          & 0.0028  & -0.1109 \\
$r_{0}$          & 0.0018  & 0.0760 \\
$\mu^{d}$      & 0.0923 & 0.0513 \\
$x$                 & 0.0999 & 0.0345 \\
$v^{L}_{0}$    & 0.0133 & 0.0266 \\
$\phi$             & 0.0082 & 0.0177 \\
$e_{0}$          & 0.0178 & -0.0072 \\
$e_{1}$          & 0.1154 & 0.0046 \\
$C_{A}$          & -0.0023 & N/A \\
$C_{L}$           & 0.0009 & N/A \\ [1ex]
\hline
\end{tabular}\\[1ex]

{\bf Table 2.} PRCC values for projected annual incidence in 2100
 in the overall population, alongside corresponding values from the
original Hill model.
\end{table}

From Table 2, we see that the PRCC values in the Extended Hill Model and the
Original Hill Model match up reasonably well.  In both cases, $\sigma^{L}$ is
the most influential parameter, with PRCC values around -0.93.  The cost
parameters $C_{A}$ and $C_{L}$ have PRCC values close to zero, which is expected
because varying the treatment costs should not affect the incidence rate of TB.
Parameters in the original Hill model with small PRCC magnitudes (less than
0.15) also have small PRCC values in the extended model, while parameters with
larger PRCC magnitudes (greater than 0.35) similarly have large PRCC values in
the extended model.  This validates the non-economic components of our model
against the original Hill model.   \\

\begin{table}[h]
\centering
\begin{tabular}{l r r r}
\hline\hline\\
Parameter & Latent Costs & Active Costs & Total Costs\\ [0.5ex]
\hline\\
$\sigma^{L}$ & 0.9612  & -0.9284 & 0.4169 \\
$v^{L}_{1}$  & -0.4190 & 0.6470  & 0.3533 \\
$f$          & 0.7467  & 0.5493  & 0.8083 \\
$p$          & 0.3371  & 0.8810  & 0.8776 \\
$ARI_{0}$    & 0.3920  & 0.6728  & 0.7337 \\
$q$          & 0.3837  & 0.6573  & 0.7200 \\
$g$          & 0.1120  & 0.5369  & 0.5182 \\
$\sigma^{F}$ & -0.1138 & -0.5631 &-0.5435 \\
$r_{1}$      & 0.1253  &  0.0658 & 0.1401 \\
$r_{0}$      & 0.1325  & 0.0878  & 0.1658 \\
$\mu_{d}$    & 0.0249  & 0.0560  & 0.0613 \\
$x$          & 0.0214  & 0.1282  & 0.1253 \\
$v^{L}_{0}$  & -0.3103 & 0.0502  &-0.1867 \\
$\phi$       & -0.0023 & -0.0102 &-0.0107 \\
$e_{0}$      & 0.0081  & 0.0253  & 0.0269 \\
$e_{1}$      & 0.0804  & 0.1653  & 0.1926 \\
$C_{A}$      & 0.0011  & 0.7024  & 0.6385 \\
$C_{L}$      & 0.8515  & 0.0013  & 0.7598 \\ [1ex]
\hline
\end{tabular}\\[1ex]

{\bf Table 3.} PRCC values for cumulative US Health Care system costs from 
Latent TB treatment, Active TB treatment, and Total treatment costs
\end{table}

In Table 3, we see that the cost parameters $C_{A}$ and $C_{L}$ are highly
correlated with Active treatment costs and Latent treatment costs respectively.
These high PRCC values are expected because there is a linear relationship
between $C_{A}$ and the cumulative Active treatment cost, and similarly for
$C_{L}$ and the cumulative Latent treatment cost.  In addition, we note that
both $C_{A}$ and $C_{L}$ are influential in the total overall cost, with PRCC
values of 0.6385 and 0.7598 respectively.  Because the PRCC value for $C_{L}$ is
greater here, despite the fact that $C_{A} > C_{L}$, we can infer that
cumulative Latent TB treatment costs are projected to be greater than cumulative
Active TB costs with these estimates for cost parameters, so $C_{L}$ is a more
influential parameter for total cumulative treatment costs.

Other variables with significant PRCC magnitudes (greater than 0.5) in Table 3
include $\sigma^{L}$, $v^{L}_{1}$, $f$, $p$, $ARI_{0}$, $q$, $g$, and
$\sigma^{F}$.  Two of these parameters, $\sigma^{L}$ and $v^{L}_{1}$, have
relatively large PRCC magnitudes for both Latent and Active Costs, but smaller
PRCC magnitudes for Total Costs.  Because their PRCC values for each of these
two outcomes is different in sign, the net change to the total cost is partially
cancelled out, so these parameters have a reduced influence on final US
healthcare system costs.  On the other hand, parameters such as $f$ and $p$ have
large positive PRCC values for both Latent Costs and Active Costs, and we
observe large positive PRCC values for Total Costs as well.\\

To gain insight about strategies to reduce the cost burden for TB in the US, we
focus on the parameters with the greatest PRCC magnitudes for Total Cost aside
from $C_{A}$ and $C_{L}$, namely $f$, $p$, $ARI_{0}$, and $q$.  All these
parameters have PRCC magnitudes above 0.7, and are highly correlated with the
total cost burden of TB borne by the US projected over the next 100 years.  Out
of these, $ARI_{0}$ is based on a historical fixed value, the Annual Risk of
Infection among the US born population in the year 2000, and therefore is
unchangeable under any intervention strategy.  The parameters $p$ (the fraction
of new infections which are acute) and $q$ (the fraction of infections
progressing to infectious disease) are variables which depend on physiological
disease dynamics, and may be altered with advances in medicine or mutations in
the bacterial strains of TB.  The parameter $f$ is the fraction of FB arrivals
with LTBI, which may vary depending on US immigration policies and medical
practices.  These results suggest that treating cases of LTBI among new FB
arrivals may be the most cost efficient intervention strategy to reduce the
disease burden for TB in the US, under the assumptions of the Hill model.  \\
\end{document}